\section{Introduction}
Stochastic programming has grown in importance in recent decades,
because it allows the modeler to accurately represent planning under
uncertainty.
With strong interest in solving such problems and in
finding more efficient solution techniques,
there has arisen a need
for a test set of stochastic programming problems.

\par
One of the most popular forms of stochastic programming problems
is the multistage stochastic linear program with recourse
(MSSLP). See Section \ref{SEC:problem-statement}, for a
precise  statement of the MSSLP problem.
While MSSLPs are growing in popularity, many of the 
applications are proprietary, and therefore the models are not 
publicly available.  Test set collections of MSSLPs exist
\cite{gassmann89,postswebsite}.  However, they need 
to be enriched with newer applications.  In fact, 
in many cases, the application associated with the existing 
test case is not known.  Also, it would 
be helpful if the original applications are described in the 
notation of the original model, and related to a 
unified notation such as in 
Section \ref{SEC:problem-statement}.

In addition, it would be desirable if the data for the test
problems is available in SMPS \cite{SMPS87}, the (emerging)
standard for specifying input to software for MSSLPs.
%%

To address the above needs, we have collected a 
group of eleven problem classes from a variety of settings.
They are all MSSLPs, but of various structures 
and sizes, with randomness occuring in different parts in different
problems.
In some cases, 
problem instances were explicitly stated in the literature.  In 
other cases, we created the problems based solely on 
the problem description in the literature, and in some 
cases, there is not yet any sample problem.  

For each model application, we present a problem description, a concise problem
statement, and, if available, a numerical example given by the model authors. 
We have attempted to stay as close to the authors' notation as possible in
these subsections.  Additionally, where feasible, we present a notational
reconciliation, which shows how to transform the notation of the problem into
that
in Section \ref{SEC:problem-statement}.

Each problem class may be used to generate one or more instances of MSSLPs.  We
have created 15 such instances.  The data for these 15 test problems, as well
as six other test problems that we did not create, are available in SMPS format
\cite{SMPS87} from the authors at
\url{http://www.uwsp.edu/math/afelt/slptestset.html}.

Also available at the web site is a chapter of problem descriptions.  Each
section covers a single problem class.  At the beginning of each section, we
give a citation to the original application, a brief description of the problem
structure, and
if applicable,
the names of the SMPS files for the associated problem instances.%  We provide
%one such section as an example in Section \ref{SEC:environ}.

It is the intention of the authors to update the classes of applications and
the test problem instances as new application areas become prominent, and to
make the information that we present for each application area, as well as SMPS
inputs for each test problem instance freely available to the stochastic
programming research community.

In that spirit, we encourage colleagues to submit new problem data
with an accompanying description.  Such submissions should include the
following:
\begin{enumerate}
\item description of the application and problem notation,
\item problem statement, in the same notation,
\item numerical example, if practical,
\item reconciliation to the notation of Section \ref{SEC:problem-statement},
\item data files in SMPS format for each instance, and
\item optimal solutions for each instance and example.
\end{enumerate}

We end this section by mentioning, that the initial motivation
for the present work came from the need to carefully test the
new polynomial interior cutting plane algorithms for
stochastic programs developed in \cite{ari96}. The
design of our test problem collection was motivated and influenced by
the elegant  work of Mor\'e et al. \cite{more90,more-averick91}
on test problem collections for several classes of (deterministic)
nonlinear optimization problems, and the impact such collections
have had on the development of software for such problem classes
\cite{more-toraldo91,more-lin99,more-benson01}. As part of testing
the algorithms developed in \cite{ari96}, we have
also developed C-routines that convert SMPS data into
a data structure suitable for implementing algorithms for
stochastic programs. 
These routines will soon be made available as open source
from the authors.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% TeX-master: "main"
%%% End: 
