\documentclass[11pt,letterpaper]{article}

%\usepackage{setspace}
%\usepackage{html}
\usepackage{url}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{theorem}
\usepackage{array}
\usepackage{epic}%used in hedgepic.tex
\usepackage{multibox}%used in hedgepic.tex
\usepackage{amsmath}
\usepackage{longtable}
\usepackage{lscape}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{color}
\usepackage{graphics}
%\usepackage{latexsym}

%\hyphenation{Mul-vey}

\includeonly{
intro,
%%
problem-statement,
%%
airlift,
forest,
electric,
hedging,
RY,
chem,
environ,
network,
cargo,
phone,
bonds
%%
}
%%
\newlength{\myscratch}
\newenvironment{lp}[1]{\setlength{\myscratch}{\arraycolsep}%
\setlength{\arraycolsep}{2pt}%
\begin{array}{#1}}%
{\end{array}%
\setlength{\arraycolsep}{\myscratch}}


\newcommand{\ds}{\displaystyle}
\newcommand{\reals}{{\mathbb R}}
\newcommand{\integers}{{\mathbb Z}}
\newcommand{\trp}{^{\mbox{\scriptsize\sf T}}}
\newcommand{\rand}[1]{\mathbf{\boldsymbol{#1}}}
\newcommand{\expect}[1][1]
        {
                {\setlength{\extrarowheight}{-6pt}
                        \begin{array}[t]{c}
                                E \\ {\scriptscriptstyle \rand{#1}}
                        \end{array}
                }
        }
\newcommand{\email}[1]{\ensuremath{\mathtt{#1}}}
\newcommand{\ajfbox}[1]{\ensuremath{\begin{array}{c}#1\end{array}}}
\newcommand{\curly}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\assign}{\ensuremath{:=}}


\title{On a new collection of stochastic linear programming test
problems}
\author{K. A. Ariyawansa\thanks{Department of
Pure and Applied Mathematics, Washington State University,
Pullman, WA 99164-3113. Email: ari@wsu.edu. The work of this
author was supported in part by the U.S. Army Research Office
under Grant DAAD19-00-1-0465}\ \ and Andrew J.
Felt\thanks{Department of Mathematics and Computing,
University of Wisconsin-Stevens Point, Stevens Point, WI 54481-3897.
Email: afelt@uwsp.edu. This author was supported in part by
Washington State University under a research assistantship
during Summer 2000.}}

%\date

\begin{document}
\maketitle
\begin{abstract}
The purpose of this paper is to introduce a new test
problem collection for stochastic linear programming
that the authors have recently begun to assemble.
While there are existing stochastic programming
test problem collections, our new collection has
three features that distinguish it from existing
collections. First, our collection is web-based with
free public access, and we intend to enrich it as new test
problems become available. Indeed, we encourage submissions
of new test problems. Second,
for each test problem class we provide a short description,
a mathematical problem statement and a notational reconciliation
to a standard format. Third, for each test problem instance
in our collection we provide numerical data in the SMPS \cite{SMPS87}
format. In a companion effort we have developed a data
structure for implementing algorithms for stochastic
linear programs, and C-routines that convert data in SMPS format
to that data structure. 
These routines will also be freely available
from the authors soon.
\end{abstract}

%problems
\include{intro}%
\include{problem-statement}%
\include{airlift}%
\include{forest}%
\include{electric}%
\include{hedging}%
\include{RY}%
\include{chem}%
\include{environ}%
\include{network}%
\include{cargo}%
\include{phone}%
\include{bonds}%

\bibliography{stochlp}
\bibliographystyle{plain}

\end{document}
